---
title: "mdp86-assignment_03"
format: html
editor: visual
---

### **Problems**

Using the [NC Births dataset](https://github.com/Bio723-class/example-datasets/raw/master/nc-births.txt), generate the following plots. For each plot use appropriate readable axis labels, colors, etc.

1.  A scatter plot showing the relationship between weeks of gestation and birth weight, with points colored by birth status (premature vs full term).

2.  A density plot showing the distribution of baby birth weights for smoking and non-smoking mothers. Use `geom_rug` to depict the marginal distributions.

3.  A boxplot comparing birth weight for premature and full term babies, with mother’s smoking status indicated by color.

4.  Draw a scatter plot showing the relationship between mothers age and fathers age. Look up how to use `ggplot2::coord_fixed()` to insure that the x- and y-axes have a fixed ratio, and modify the plot limits to include the origin. Color the points by birth status.

### Load data

```{r}
library(tidyverse)
nc_births <- read_tsv("~/Bio724/nc-births.txt")
```

### 1.

```{r}
ggplot(data = nc_births, mapping = aes(x = weeks, y = weight, color = premature)) +
  geom_point()
```

### 2.

```{r}
ggplot(data = nc_births, mapping = aes(x = weight, color = smoke)) +
  geom_density(alpha = 0.5) +
  geom_rug()
```

### 3.

boxplot comparing birth weight for premature and full term babies, with mother’s smoking status indicated by color.

```{r}
ggplot(data = nc_births, mapping = aes(x = premature, y = weight, color = smoke)) +
  geom_boxplot() + 
  theme_minimal()

ggplot(data = nc_births, mapping = aes(x = smoke, y = weight, color = smoke)) +
  geom_boxplot() +
  facet_wrap(vars(premature)) +
  theme_bw()
```

### 4.

Draw a scatter plot showing the relationship between mothers age and fathers age. Look up how to use `ggplot2::coord_fixed()` to insure that the x- and y-axes have a fixed ratio, and modify the plot limits to include the origin. Color the points by birth status.

```{r}
ggplot(data = nc_births, mapping = aes(x = fAge, y = mAge)) + 
  geom_point() +
   coord_fixed(ratio = 1) +
  xlim(0, 50) +
  ylim(0, 50) +
  coord_fixed(ratio = 1)

nc_births_na_dropped <- nc_births %>% drop_na()
ggplot(data = nc_births_na_dropped, mapping = aes(x = fAge, y = mAge)) + 
  geom_point() +
   coord_fixed(ratio = 1) +
  xlim(0, 50) +
  ylim(0, 50) +
  theme_minimal() + 
  ylab("Mother's age") +
  xlab("Father's age")
```

# Data lunch assignment

I found the approach for QTL mapping overtime interesting to be able to better understand developmental changes and responses to changing environmental conditions overtime.
